---
title: "P8105 Homework 1"
author: "Grace Santos"
date: "09-18-23"
output: github_document
---

This document will involve all problems associated with HW 1 for Data Science (P8105).

## Problem 1 

The following problem focuses on the use of inline R code, plotting and the behavior of `ggplot` for different variables. 

```{r initial library and data}
#Loading the `moderndive` library and the `early_january_weather` datasets

library(moderndive)

data("early_january_weather")
```

A description of the `early_january_weather` dataset can be found below: 

To first identify the number of variables contained within this dataset, I performed the following function:

```{r variable}
ncol(early_january_weather)
```

From this, we now know that this dataset includes a total of 15 variables. They are entitled: `r colnames(early_january_weather)`. 

Next, the total number of observations were found using the function below:

```{r observation}
nrow(early_january_weather)
```

This result shows that there are 358 total observations within the data set.

```{r, library}
#To continue with my data observation analysis, I loaded the tidyverse library in order to pull information about different variables and eventually plot the given data as well. 
library(tidyverse)
```

By taking a closer look at the data, we can determine the following information about the provided data set variables: 

* For all data observations, the origin, year and month variables all have the same values. This implies that each collected data point was gathered from an origin point that can be abbreviated as 'EWR', during the month of January, in the year 2013. 
* Later on in this problem, the variables for temperature and for the recorded time (in hours) will be used to generate a plot. The temperatures ranged from `r range(pull(early_january_weather, temp))` degrees.
* Also, there is data provided for wind, precipitation, atmospheric pressure and other meteorological statistics included in this dataset. Upon closer inspection, I did notice that for many of the wind data observations, there are instances of missing or zero-valued data points within these variable categories.

**Overall, the mean temperature  was `r mean(pull(early_january_weather, temp))` degrees.**

A scatterplot of temperature vs. time is shown below. A pdf version of the generated plot will be saved within the project folder. 

```{r plot}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()+ ggtitle("Temperature vs. Time in Hours")

ggsave("hw1_scatterplot.pdf")
```

By looking at the generated scatterplot, the data shows that over the first ~2 weeks of January in 2013, the temperature trended upwards and the humidity fluctuated with temperature. The peak in humidity was found during the latter few days of data collection as temperatures began to regulate before the final peak in temperature. 

## Problem 2 

Creating a data frame from instructions provided. 
```{r, create df}
problem2_df = tibble(
  sample = rnorm(10),
  vec_log = any(sample > 0),
  vec_char = vector(mode = "character", length = 10),
  vec_factor = factor(c("sun", "star", "moon", "star", "star", "sun", "moon", "moon", "star", "sun"), levels = c("sun", "star", "moon"))
)
```

Attempting to take the mean of each variable in the dataframe. 
```{r, mean df}
levels(problem2_df)
```

Convert variables from one type to the other using `as.numeric` function. 
```{r, convert var}

```

