---
title: "P8105 Homework 1"
author: "Grace Santos"
date: "09-18-23"
output: github_document
---

This document will involve all problems associated with HW 1 for Data Science (P8105).

## Problem 1 

The following problem focuses on the use of inline R code, plotting and the behavior of `ggplot` for different variables. 

```{r initial library and data}
library(moderndive)
library(tidyverse)

data("early_january_weather")
```

A description of the `early_january_weather` dataset can be found below: 

To first identify the number of variables contained within this dataset, I performed the following function:

```{r variable}
ncol(early_january_weather)
```

From this, we now know that this dataset includes a total of 15 variables. They are entitled: `r colnames(early_january_weather)`. 

Next, the total number of observations were found using the function below:

```{r observation}
nrow(early_january_weather)
```

This result shows that there are 358 total observations within the data set.

```{r isolating var, echo = FALSE}
temp = pull(early_january_weather, temp)
rain = pull (early_january_weather, precip)
wind_speed = pull (early_january_weather,wind_speed)
```

By taking a closer look at the data, we can determine that in January of 2013, the temperatures ranged from `r range(temp)` degrees, the maximum precipitation was measured as `r max(rain)` and the windspeed averaged `r mean(wind_speed)`.

Overall, the mean temperature  was `r mean(temp)` degrees. 

A scatterplot of temperature vs. time is shown below. A pdf version of the generated plot will be saved within the project folder. 

```{r plot}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()+ ggtitle("Temperature vs. Time in Hours")

ggsave("hw1_scatterplot.pdf")
```

