---
title: "P8105 Homework 1"
author: "Grace Santos"
date: "09-18-23"
output: github_document
---

This document will involve all problems associated with HW 1 for Data Science (P8105).

## Problem 1 

The following problem focuses on the use of inline R code, plotting and the behavior of `ggplot` for different variables. 

```{r initial library and data}
#Loading the `moderndive` library and the `early_january_weather` datasets

library(moderndive)

data("early_january_weather")
```

A description of the `early_january_weather` dataset can be found below: 

To first identify the number of variables contained within this dataset, I performed the following function:

```{r variable}
ncol(early_january_weather)
```

From this, we now know that this dataset includes a total of 15 variables. They are entitled: `r colnames(early_january_weather)`. 

Next, the total number of observations were found using the function below:

```{r observation}
nrow(early_january_weather)
```

This result shows that there are 358 total observations within the data set.

```{r, library}
#To continue with my data observation analysis, I loaded the tidyverse library in order to pull information about different variables and eventually plot the given data as well. 
library(tidyverse)
```

By taking a closer look at the data, we can determine the following information about the provided data set variables: 

* For all data observations, the origin, year and month variables all have the same values. This implies that each collected data point was gathered from an origin point that can be abbreviated as 'EWR', during the month of January, in the year 2013. 
* Later on in this problem, the variables for temperature and for the recorded time (in hours) will be used to generate a plot. The temperatures ranged from `r range(pull(early_january_weather, temp))` degrees.
* Also, there is data provided for wind, precipitation, atmospheric pressure and other meteorological statistics included in this dataset. Upon closer inspection, I did notice that for many of the wind data observations, there are instances of missing or zero-valued data points within these variable categories.

**Overall, the mean temperature  was `r mean(pull(early_january_weather, temp))` degrees.**

A scatterplot of temperature vs. time is shown below. A pdf version of the generated plot will be saved within the project folder. 

```{r plot}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()+ ggtitle("Temperature vs. Time in Hours")

ggsave("hw1_scatterplot.pdf")
```

By looking at the generated scatterplot, the data shows that over the first ~2 weeks of January in 2013, the temperature trended upwards and the humidity fluctuated with temperature. The peak in humidity was found during the latter few days of data collection as temperatures began to regulate before the final peak in temperature. 

## Problem 2 

Creating a data frame from instructions provided. 
```{r, create df}
problem2_df = tibble(
  sample = rnorm(10),
  vec_log = sample > 0,
  vec_char = c("sun", "moon", "star", "sun", "sun", "star", "moon", "sun", "star", "moon"),
  vec_factor = factor(vec_char) 
)
```

Attempting to take the mean of each variable in the dataframe. 
```{r, mean df}
mean(pull(problem2_df, sample))
mean(pull(problem2_df, vec_log))
mean(pull(problem2_df, vec_char))
mean(pull(problem2_df, vec_factor))
```
The provided results show that the mean was able to be calculated for the first two variables in my dataframe (the random sample and logical vector) and did not work for the last two vectors in my dataframe (the character vector and the factor vector). The mean of the character and factor vectors were not able to be calculated because they are neither numerical or logical. 


Converting variables from one type to the other using `as.numeric` function for the logical, character and factor variables. 
```{r, convert var, eval = FALSE}
num_vec_log = as.numeric(pull(problem2_df, vec_log))
num_vec_char = as.numeric(pull(problem2_df, vec_char))
num_vec_factor = as.numeric(pull(problem2_df, vec_factor))
```
These results show that the logical vector and factor vector were able to be converted into a numeric vectors and the character vector was not. The factor vector was able to be transformed into a numerical vector because factors are stored as integers, thus the `as.numeric` function is able to turn change the classification of these integers from factors to numerical entries. Before it was turned into a numerical vector, the factor vector could not be evaluated for the mean, but after converting it, the mean can be taken. The character vector was not able to be transformed into a numerical vector and resulted in `NA` for all values since character vectors are stored as words and therefore can not be transformed into numbers. This explains why the mean cannot be taken for a character vector both before and after converting it to numeric. 
